---
title: "| Environmental and Development Economics\n| Module 7 - Climate and Migration
  \n"
author: |
  | Raahil Madhok
  | UMN Applied Economics
date: "`r Sys.Date()`"
output:
  beamer_presentation:
    theme: Montpellier
  slidy_presentation: default
  ioslides_presentation: default
classoption: aspectratio=169
header-includes:
- "\\AtBeginDocument{\\title[Module 7 - Climate and Migration]{Environmental and Development
  Economics \\\\ Module 7 - Climate and Migration}}"
- \input{header.tex}
- \usepackage{gensymb}
---


## 
\centering
\Huge Lecture 12

\Large Climate and Migration

## Housekeeping

- APEC Seminar Tomorrow: Stephanie Onder
\vfill
- Peer review due at midnight
  - I will return tomorrow
\vfill
- Proposal presentations start next week
  - **Oct 15:** 1) Matt B 2) Jovin L 3) Ryan M
  - **Oct 17:** 1) Gustavo M 2) Giang T 3) Sarah W
\vfill
- Office hours will continue until proposal deadline

## Replication Assignment (Due Oct 21st)

- Choose any envirodevonomics paper with data/code files
  - Must be published in either 1) general econ or 2) enviro/dev field journal 
  
- Replicate at least 2 main tables/figures
  - Output everything to latex

- Write your own code and comment it!

- Extension
  - Must be "carefully thought out" and yield new insight
  - e.g. heterogeneity, non-linearity, robustness check
  
- Write-up (6 pages doubled spaced) 
  - summary, replication process, results, extension
  - Compare your findings with original paper. Interpret.

##
\centering
\huge Climate and Migration

## Environmental disasters are unequal 

- Cross-country heterogeneity
  - Inequality in exposure (rich vs. poor countries)
\vfill
- Within-country heterogeneity
  - Coastal regions hit harder by floods
  - Agricultural regions hit harder by droughts
  - Biodiverse regions hit harder by rainfall shocks
\vfill
- Sectoral heterogeneity
  - Weather-dependent agriculture vs. services
\vfill

**Implication $\longrightarrow$ migration!**

## Unequal Climate Impacts

Historic average temperature (1986-2005)
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "95%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/climate_fig1.png",
                        auto_pdf = TRUE
                        )

```

## How do we study flows of people in response to climate change?

- Impacted region (push factors)
  - Direct effect: environmental extremes push people out
  - Indirect effect: violence, famine, etc. push people out
\vfill
- Receiving region (pull factors)
  - Direct effect: less exposure to environmental extremes draws migrants (wage impacts)
  - Indirect effect: e.g. housing prices
\vfill
- Most research: environment $\leftrightarrow$ migration in partial equilibrium
  - Elasticity of migration w.r.t. environmental quality
  - Elasticity of wages w.r.t environmental quality
\vfill
- **New frontier: general equilibrium!**

## Existing Evidence 
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/litreview.png",
                        auto_pdf = TRUE
                        )

```

## Spatial Models of Migration

- Economists are recently thinking about spatial equilibrium
\vfill
- Migrants choose where to live based on costs and benefits
  - Not only wage considerations
  - Amenity values, **exposure to environmental extremes**
\vfill
- When there is an environmental shock to the system...
  - Who moves?
  - What adjustments take place? Housing, land, wages, etc.
  - What is the new spatial equilibrium?
  - What happens to inequality?
\vfill
- This sets the framework for studying environment and migration
  
## Model Set Up

- Individual $i$ of wage group $j$ choose location $k$ to maximize utility:

$$ \max_{k} \quad \underbrace{v_{jk}}_\text{observed} + \underbrace{\epsilon_{ijk}}_\text{unobserved}$$

- The econometrician observes representative utility:

$$ v_{jk} = \alpha_{j}p_{k} + \beta e_{k} + x_{k}\gamma + \delta_{jk}$$

- housing price $p_{k}$
- environmental "bad" $e_{k}$ (flood risk, extreme temperature)
- observed amenities $x_{k}$, unobserved amenities $\delta_{jk}$
- price elasticity can differ by wage group, $\alpha_{j}$

## Model Set Up

- Logit probability ($\epsilon_{ijk} \sim EV$) yields migration probability (see module 4 slides):

$$ \pi_{ijk} = \pi_{jk} = \frac{e^{v_{jk}}}{\sum_{l}e^{v_{jl}}}$$

- since individuals w/n wage group $j$ have common choice probabilities

- In equilibrium, prices $p = p_{k}$ clear housing markets

$$ n_{k}^{D}(p^{*}) = n_{k}^{S}(p^{*}) \quad \forall k \quad \text{where} \quad n_{k}^{D} = \sum_{i} \pi_{ijk}, \quad n_{k}^{S} = \bar{n}_{k} $$

\vfill
- **Migration arises from wage-specific price elasticities and endogenous prices**
  - Environmental shocks $\uparrow$ inequality. Why?

## Estimation

- Goal: recover $\alpha_{j}$, $\beta$, $\gamma$
  - Tells us equilibrium prices and choice probabilities for any pattern of $e_{k}$
\vfill
- Inverting migration probabilty yields:
\begin{align*}
ln \pi_{jk} - ln \pi_{jo} & = v_{jk} - v_{j0} \\
  & = \alpha_{j} \Delta p_{k} + \beta \Delta e_{k} + \Delta x_{k} \gamma + \Delta \delta_{jk}
\end{align*}
\vfill

- **Challenge:** prices are endogenous i.e. correlated with unobserved amenities!
  - high-amenity places attract skilled migrants who bid up prices
  - Solution: cost shifting instrument

## Hisao (2024): Sea Level Rise and Urban Inequality in Indonesia

- Flood exposure: grid cells $<$ sea level for 1,3, 5m sea level rise (hydrological model)
\vfill
- Price instrument: ruggedness (housing cost supply shifter)
  - Exclusion restriction: terrain shifts housing supply but not demand
\vfill
- Define object of interest for simulation: flood exposure for wage group $j$: 
$$F_{j} = \sum_{k} f_{k}\pi_{jk}$$
- And price incidence by wage group $j$: 
$$P_{j} = \sum_{k}p_{k}\pi_{jk}$$

## Impact of flooding on inequality
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/hsiang2024_fig1.png",
                        auto_pdf = TRUE
                        )

```

- Current flooding: poor people's flood exposure is 1.4x that of rich people
- 5m sea level rise: inequality of exposure doubles
- Prices: richer people move, lower prices compensate for higher flood exposure

## Is inequality driven by prices or migration?
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "90%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/hsiang2024_fig1.png",
                        auto_pdf = TRUE
                        )

```
- Evaluate $F_{j}$ with projected flooding and current $\pi_{jk}$ (no migration)
- Evaluation $P_{j}$ with counterfactual prices and current $\pi_{jk}$ (no migration)
- Reveals direct impact of floods on rich/poor **without migration**
- No $\Delta$ in inequality in this scenario i.e. **migration drives unequal exposure!**

## Reduced Form Literature

- **Impact of Migration on Environment**
  - \textcolor{red}{Brewer et al. (2022, Uganda)}; Hornbeck and Naidu (2014, USA); Rozelle et al. (1991, China); Salemi (2021, Africa)}
\vfill
- **Impact of Environment on Migration**
  - Morten (2019, India); \textcolor{red}{Mueller et al. (many papers)}; Missrian and Schlenker (2017, dev countries); Chen et al. (2022, China)
\vfill
- **Migration as Adaptation to Environmental Extremes**
  - \textcolor{red}{Cecato et al. (2024, Indonesia)}; Blakeslee et al. (2020, India)
\vfill
- **Migration and Environment in General Equilibrium**
  - \textcolor{red}{Madhok et al. (2024, India)}; Asher et al. (2024, India); Blakeslee et al. (2022, India)

## Impact of Migration on the Environment: Brewer et al. (2022)

- **Research Question:** What is the impact of rural-urban migration on land use?
  - What are the intensive vs. extensive margin impacts?
\vfill
- **Setting:** Uganda from 2005-2019 
  - 70\% of land is agricultural
\vfill
- **Data:** Many rounds of the LSMS survey (3200 households)
\vfill
- **Research Design:** SSIV: urban consumption (shift) x migrant network (share)
\vfill
- **Results:** Migration causes agricultural contraction and specialization
  - Leads to a reduction in tree cover loss

## Conceptual Framework
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "70%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/brewer2022_fig1.png",
                        auto_pdf = TRUE
                        )

```

## Estimation

- 1st stage: $IHS(M_{ijt}) = \underbrace{\beta_{11}Z_{jt}M_{ij0}}_\text{shift-share IV} + \beta_{21}Z_{jt} + \beta_{31}W_{jst}+\alpha_{1i} + \psi_{1w} + \gamma_{1t} + \epsilon_{1ijt}$
\vfill
- 2nd stage: $IHS(Y_{ijt}) = \beta_{12}(\hat{M}_{ijt}) + \beta_{22}Z_{jt} + \beta_{32}W_{jst}+\alpha_{2i} + \psi_{2w} + \gamma_{2t} + \epsilon_{2ijt}$
\vfill
- Outcome: $M_{ijt}$ is number of months that hh member $i$ was absent in year $t$
\vfill
- Shift-share IV: 
  - $Z_{jt} = \sum_{d=1}^{n}(\frac{C_{dt}}{dist_{jd}})$
  - $C_{dt}$ is mean consumption in district d
  - $Z_{jt}M_{ij0}$ is spatially weighted consumption shocks

## Results: Agricultural extent
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "90%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/brewer2022_fig2.png",
                        auto_pdf = TRUE
                        )

```
- rural-urban migration leads to agricultural contraction (extensive margin)
- rural-urban migration has no impact on area planted for annual vs. perennial crops

## Results: Crop Specialization
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "90%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/brewer2022_fig3.png",
                        auto_pdf = TRUE
                        )

```
- migration causes households to specialize
- 1) reduce number of different crops, 2) invest proportionally more in fewer crops

## Results: higher district level tree cover
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/brewer2022_fig4.png",
                        auto_pdf = TRUE
                        )

```

## Impact of Environment on Migration: Chen and Mueller (2018)

- **Research Question:** How do farmers adapt to sea-level rise? 
  - At what point do farmers migrate?

\vfill
- **Setting:** Coastal Bangladesh, 2003-2011
\vfill
- **Data:** Several sources
  - Migration: Vital registration system
  - Agriculture: Household Income and Expenditure Surveys
  - Floods: NASA MODIS water index
\vfill
- **Research Design:** Fixed effects regression
\vfill
- **Results:** Indundation *does not* trigger migration
  - Migration is driven by adverse effects of soil salinity on agriculture


## Conceptual Framework

- Recent studies show minimal migratory response to flooding
  - Monsoons lead to temporary displacement, but people return
  - Delayed yield benefits
  - Many farmers have adapted ag practices to accommodate inundation
\vfill
- Inundation may not be the right ``push'' factor
  - What exactly is triggering flood-related migration?
\vfill
- Hypothesis: gradual increases in salinity (not flooding) will bear modest effects on
migration until less costly adaptation strategies in the agricultural
sector have been exhausted

## Empirical Strategy
- Impact of salinity and flooding on agricultural incomes: 

$$ Y_{hjt} = \theta_{F} F_{jt-1} + \theta_{S}S_{jt-1} + \beta Z_{hjt} + \delta X_{hjt-1} + \beta_{S}^{88}S_{j,1988} + \tau_{t} + \mu_{hjt} $$

- $Y_{hjt}$ = annual crop revenue for household $h$ in subdistrict $j$ at time $t$
- $F_{jt-1}$ = lagged flooding; $S_{jt-1}$ = lagged salinity
- $Z_{hjt}$ = demographic and wealth covariates
- $X_{hjt-1}$ = lagged environmental exposure (rain, temp, etc.)
\vfill
- Impact of salinity and flooding on migration

$$M_{hjt} = \beta_{F} F_{jt-1} + \beta_{S}S_{jt-1} + \alpha Z_{hjt} + \gamma X_{hjt-1} + \beta_{S}^{88}S_{j,1988} + \rho_{t} + \epsilon_{hjt}$$

## Results: Effect of flooding/salinity on farm revenue
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/mueller2018_fig1.png",
                        auto_pdf = TRUE
                        )

```

- 1$\sigma$ increase in salinity reduces crop revenue by 0.1$\sigma$ (2500 taka)
- No impact on total farm revenue
- hh shift to aquaculture to mediate income risk as salinity compromises agriculture 

## Results: Effect of flooding/salinity on migration
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/mueller2018_fig2.png",
                        auto_pdf = TRUE
                        )

```
- 1$\sigma$ increase in soil salinity increases internal migration by 0.43pp
- 1$\sigma$ increase in soil salinity decreases international migration by 0.19pp
- International migration and aquaculture may be substitutes

## Migration as Adaptation: Cecato et al. (2024)

- **Research Question:** Can migration mitigate weather damages?

\vfill
- **Setting:** Indonesia over 20 years
\vfill
- **Data:** 
  - Migration: 32,000 individuals from IFLS survey
  - Weather: Gridded rain/temperature 
\vfill
- **Research Design:** Shift-share IV
  - Shift: rainfall shock
  - Share: ethnic networks
\vfill
- **Results:** 1$\sigma$ $\downarrow$ in precipitation at origin $\downarrow$ non-migrant consumption 
  - Ability to migrate leads to 14\% increase in consumption over medium run

## Three facts

- There is high migration in Indonesia
  - 4.3\% percent of population per year migrates internally
\vfill
- Rainfall is an important determinant of migration in Indonesia
  - Kleemans and Magruder (2018)
\vfill
- Mobility is associated with ethnicity 
  - Propensity to migrate persists over time within groups (Auwalin 2020)

## Estimation

- **First stage regressions:**

$$M_{ikt} = a_{0}M_{ikt} + a_{1}(Networks_{ik} \cdot R_{kt}) + a_{2}R_{kt} + \phi_{i} + \phi_{t} + X_{ikt} + \epsilon_{ikt}$$

$$M_{ikt} \times R_{kt} = b_{0}M_{ikt} + b_{1}(Networks_{ik} \cdot R_{kt}) + b_{2}R_{kt} + \phi_{i} + \phi_{t} + X_{ikt} + \epsilon_{ikt}$$

\vfill

- **Second stage:**
$$Y_{ikt} + \beta_{0}\hat{M}_{ikt} + \beta_{1}(\widehat{M_{ikt} \times R_{kt}}) + \alpha_{1}R_{kt} + \phi_{i} + \phi_{t} + X_{ikt} + \epsilon_{ikt}$$

\vfill

- $Y_{ikt}$: log consumption of hh $i$ in county $k$ in year $t$
- $R_{kt}$: rainfall shock (z-score)
- $Networks_{ik}$: share of ppl of same ethnicity in origin at age 12 who moved away

## Results: Migration mitigates weather damage
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/cecato2024_fig1.png",
                        auto_pdf = TRUE
                        )

```
- 1$\sigma$ decrease in rainfall reduces consumption by 1.82\%
- Ability to migrate more than offsets this

## Migration and Agriculture in General Equilibrium: Madhok et al. (2024)

- **Research Question:** How does agriculture unfold when labor exits?
\vfill
- **Challenge:** Two competing channels: labour (direct/PE) + prices (indirect/GE)
\vfill
- **Setting:** India --- very high labour misallocation
\vfill
- **Data:** IHDS-I and IHDS-II panel surveys (42,000 households)
\vfill
- **Research Design:** Shift-share IV
  - Shift: destination income shocks
  - Share: 1) inverse distance, 2) number of males at home
\vfill
- **Results:** migration $\downarrow$ agriculture near cities, but $\uparrow$ in remote areas
  - i.e. spatial reorganization of agriculture

## Partial Equilibrium

- **Shift share IV:** 
\begin{equation*}
	z_{idt} = \underbrace{\varphi_{i}}_\text{productivity (share I)} \times \underbrace{\sum_{d' \in \Theta/ d} \frac{1}{\tau_{dd'}}}_\text{inverse distance (share II)} \times \quad \underbrace{(inc_{d't} \cdot pop_{d'})}_\text{shift}
\end{equation*}

- **2SLS Equations:**

\begin{align*}
   & M^{labor}_{idt}  = \mu_{1}z_{idt} + \mu_{2}s_{dt} + \mu_{3} inc_{dt} + \Gamma X'_{dt} + \alpha_{i} + \gamma_{t} + \varepsilon_{ijdt} \\
    & Y_{idt}  = \beta_{1}\widehat{M^{labor}_{idt}}
+ \beta_{2}s_{dt} + \beta_{3}inc_{dt} + \Gamma X'_{dt} + \alpha_{i} + \gamma_{t} + \eta_{ijdt}
\end{align*}

## PE Results: No labor-capital substitution
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/madhok2024_fig1.png",
                        auto_pdf = TRUE
                        )

```

## Spatial Equilibrium: Visual Evidence
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/madhok2024_fig2.png",
                        auto_pdf = TRUE
                        )

```

## Migration and Market Adjustments
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "80%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/madhok2024_fig3.png",
                        auto_pdf = TRUE
                        )

```
- Mass migration from the crop market increases crop prices
- Mass migration from the land market lowers land prices 

## Migration and the Spatial Reorganization of Agriculture
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "75%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/madhok2024_fig4.png",
                        auto_pdf = TRUE
                        )

```
- PE (agricultural contraction) dominates in peri-urban areas
- GE (agricultural expansion) dominates in remote areas
- Result: food shortage induced by direct channel compensated by markets

##
\centering
\huge New Directions for Environment and Development Economics 

\Large Open Research Areas

## Exciting Area for Research

- **Main takeaway from course:**
  - Many reasons for poor environmental quality in LMICs
  - Evidence on many topics is very thin!

\vfill
- Barriers to research are falling
  - Remote sensing (e.g. cashews in West Africa)
  - Administrative data (e.g. Annual Survey of Industries)
  - Poverty proxies (e.g. nightlights, other geospatial proxies)
\vfill

## Much Room for Methodological Innovations

- I have tried to emphasize **applied theory**
  - We can take existing models and adapt them to LMIC contexts
  - E.g. add credit constraints, factor market constraints
  - Many new insights we can take to the field!
\vfill
- Other methodological innovations
  - Demand estimation (e.g. Ito and Zhang, 2020; Kremer et al., 2013)
  - Field experiments (e.g. Saavedra, 2024; Duflo et al., 2013)
  - Lab-in-the-field (e.g. Dean and Berkouwer, 2022)
  - Reduced form quasi-experimental

## Innovation in Theory

1. Environmental and development econ study market failures
    - New theory can help explain interactions
    - e.g. Burgess et al. (2012), Jack et al. (2024), Madhok and Gulati (2024), Braganca and Dahis (2022)
\vfill   
2. IO and behavioural economics have clear intersections w/ envirodevonomics
    - WTP elicitation (revealed pref, stated pref, etc.)
    - e.g. Ito and Zhang (2020), Berkhouwer and Dean (2022), Kremer et al. (2013)
    - See reading list for many other WTP for environment papers
\vfill
3. [New] Macro-structural and trade 

## Research Frontiers: $MWTP_{e}$

- What is $MWTP_{e}$ in poor countries?
  - Lots of work on pollution (air/water)
  - Less work on biodiversity/forests
\vfill
- What is the role of market failures in explaining low $MWTP_{e}$?
  - Credit markets, labor markets, property rights, information access
\vfill
- Is low state capacity reflected in low MWTP?

## Research Frontiers: Environmental policy design in LMICs
 
- What determines successful environmental policy in LMICs?
  - What is role of monitoring, enforcement, information asymmetries?
  - New areas: e.g. trust (Jack et al., 2024)
\vfill
- Does information provision to public change exposure to environmental risk?
\vfill
- How can environmental policy be tailored to accommodate informal sector?

## Research Frontiers: Political Economy

- How do political structures lead to rent-seeking and overexploitation?
  - Can conservation policy mitigate rent-seeking behavior?
\vfill
- How does leader identity influence environmental outcomes?
  - Ethnic alignment (e.g. Madhok and Gulati, 2024)
  - Gender (e.g. Jagnani and Mahadevan, 2024)
  - Minority/majority: e.g. (Gulzaar et al, 2023)
\vfill
- Do extractive institutions cast a "long shadow" on the environment?

##
\centering
\huge Thank you!

\large Next week: Proposal Presentations
