---
title: "| Environmental and Development Economics\n| Module 3 - How does the environment affect development?
  \n"
author: |
  | Raahil Madhok
  | UMN Applied Economics
date: "`r Sys.Date()`"
output:
  beamer_presentation:
    theme: Montpellier
  slidy_presentation: default
  ioslides_presentation: default
classoption: aspectratio=169
header-includes:
- "\\AtBeginDocument{\\title[Module 3 - environment and development]{Environmental and Development
  Economics \\\\ Module 3 - How does the environment affect development?}}"
- \input{header.tex}
- \usepackage{gensymb}
---


## 
\centering
\Huge Lecture 4

\large How does the environment affect economic development?

## Housekeeping

- Consider pitching your research idea to me during office hours
\vfill
- Next week, you have TWO options for the presentation paper
\vfill
- Your class presentation grade is an average of each one
\vfill
- Question about revealed preference WTP lecture (Sept 19)

## Today

- **Guiding question:** how does the environment affect economic development?
\vfill
- Start identifying envirodevonomics parameters
\vfill
- Today's focus: health effects
\vfill
- Next time: productivity and direct income effects

## Recall our model of environment and development
\begin{align}
MWTP_{e} = \frac{\lambda_{e}}{\lambda_{y}} 
 = \frac{1}{\lambda_{y}} \left(\frac{\partial u}{\partial e} + \frac{\partial u}{\partial h}{\color{red} \frac{\partial h}{\partial e}}\right) + \frac{\partial \Delta y}{\partial e} + \frac{\partial \Delta y}{\partial h}{\color{red}\frac{\partial h}{\partial e}} \notag
\end{align}

- Most research in environment/development economics seeks to identify $\frac{\partial h}{\partial e}$
  - At this point, $\frac{\partial h}{\partial e}>0$ is unambiguous
  - Focus is on **magnitudes**
  - Less work on mechanisms and moderators
\vfill
- What about indirect effects via productivity and health ($\frac{\partial \Delta y}{\partial h}\frac{\partial h}{\partial e}$)? 
  - This is the research frontier
  - Next lecture

## Existing Work to Date

- Largely from developing countries (see Janet Currie papers)
  - Review paper: ``What Do We Know About Short- and Long-Term Effects of Early-Life Exposure to Pollution?''(Currie et al. 2014)
\vfill
- Less in developing countries
  - Most benchmark magnitudes: Bharadwaj and Eberhard (2008); Jayachandran (2009); Chen et al. (2013); Greenstone and Hanna (2014); Arceo et al. (2016); Cesur et al. (2016); He et al. (2016); Barwick et al. (2018); Heft-Neal et al. (2018); Chang et al. (2019); Heft-Neal et al. (2019); Pullabhotla (2019); Bombardini and Li (2020); Fan et al. (2020); He et al. (2020); Adhvaryu et al. (Forthcoming); Madhok et al. (2024, Working Paper)
\vfill
- Less work on non-health effects
  - See Aguilar-Gomez et al. (2022) for review
\vfill
- Empirical innovations: new quasi-experiments for identifying pollution impacts
  - Thermal inversion, wind direction IV (see Oliva \& Deryugina)
  - Newest technique: HYSPLIT dispersion model (NOAA)
  
## Existing Work to Date

- Consensus that there are large negative effects in developing countries
  - Little evidence on why
\vfill
- Open questions:
  - Are marginal damages higher or lower in LMICs? (Colmer et al.,2021)
  - Are differences in marginal damages because of non-linear dose-response
  - Are differences in marginal damages because of weak instututions/development?
  - Long term effects of exposure?
\vfill
- Conceptual frameworks needed

## Lets formalize our intuition
- Weather can harm health directly (physiological stress)
  - Or indirectly, e.g. by lowering incomes of those in weather-dependent (outdoor) jobs

- There is a market for defensive goods (e.g. medicine, AC, etc.) which $\uparrow$ P(survival)

- But in developing countries, credit constraints restrict consumption smoothing

- Lets incorporate this into a model and study the comparative statics
  - Goal: explain how weather affects health, considering financial constraints
  - Health is endogenously determined by consumption and health investments

## Setup
- **Agents:** value consumption $c_{it}$ and health $h_{it}$.
- **Objective:** Maximize lifetime utility
$$\mathbb{E}_{0} \left[\sum_{t=0}^{\infty} \beta^{t}u_{i}(c_{it}, h_{it}) \right]$$
\vfill
- **Health Status**: Determined by health investments ($q_{it}$) + exogenous weather ($z_{it}$)
$$h_{it} = h_{i}(\underbrace{q_{it}}_{(+)}, \underbrace{z_{it}}_{(-)})$$

## Setup

**Production**

- $c_{it}$ produced by $f(\cdot)$, which requires capital + labor
- Assume defensive good $q_{it}$ traded at price $P_{t}$ (numeraire)

**Labor**

- Agent endowed with labor $a_{it}$ (random), sold at wage $\omega_{t}$
- Labor can be affected by weather ($z_{it}$)
$$a_{it} = a_{i}(z_{it})$$

## Financial Market

- Agent can (partially) save/borrow to (partially) smooth consumption over $t$

- **Financial Constraints**: Savings ($x_{it}$) and borrowing are limited by:
  - Borrowing constraint (mininum savings): $x_{it} \geq -b_i(r_t)$
  - Saving constraint (maximum savings): $x_{it} \leq b_i(r_t)$
  - $b(\cdot)$ unrestricted s.t. agent can be unconstrained
\vfill

- **Budget Constraint**: income is used for consumption and health investments:
  $$x_{i,t+1} \leq r_t x_{it} + w_t a_i(z_{it}) - P_t q_{it} - c_{it}$$


## Equilibrium

- All markets must clear each $t$, which endogenously determines prices $(r_{t}, w_{t})$
  - Note $P_{t}$ is exogenously set in world market
\vfill 
- Returns to capital/labor are functions of their aggregate
  - Capital demand: from production function $f(\cdot)$
  - Capital supply: savings transformed into capital for production 
\vfill
- **Market Equilibrium**: set optimal demand for labor/capital = aggregate supply  
  - where $K_{t} = \int x_{it}di$
  - wages $w_t$ and interest rates $r_t$ are determined in equilibrium as:
  $$w_t = w(K_t), \quad r_t = r(K_t) $$

## Mechanisms relating weather to health

- **Direct health effect:** random shock $z_{it}$ enters health directly: $h_{it} = h_{i}(q_{it}, z_{it})$
  - Physiological channel: extreme temperature stresses cadiovascular/respiratory system
  - Impact can depend on region, $h_{i}(\cdot)$
\vfill
- **Indirect health effect:** through productivity $a_{i}(z_{it})$
  - Likely in rural areas, where ag workers are outside
\vfill  
- **Mitigation:** agent minimizes neg. utility shock from $z_{it}$ by purchasing $q_{it}$.
  - This could also be opp cost of working inside (at lower $w_{it}$) to avoid outdoors
  - Effect of mitigation limited by $\frac{\partial h_{i}(q_{it}, z_{it})}{\partial q_{it}}$
\vfill
- **Smoothing:** agent can also mitigate by reallocating resources b/w periods
  - Agents with near-binding borrowing constraints experience worse shocks

## Comparative Statics
- **Object of interest**: aggregate covariance b/w health \( h_{it} \) and weather \( z_{it} \):
    $$H_t = \left| \frac{\int h_{it} z_{it}^{(T)} di}{\int h_{it} di} \right|$$
\vfill
- **Rural Areas**: Consider a set of changes that:
  - $\uparrow$ responsiveness of $a_{i}(z_{it})$ to $z_{it}^{T}$ (more exposed to weather extremes).
  - $\uparrow$ prices for health goods $P_t$ above that of urban (lower market access)
  - $\downarrow$ incomes by reducing $a_{i}(z_{it})$ for any $z_{it}$
  - Any one of these leads to increase in $H_{t}$
\vfill
**Prediction:** rural areas will have stronger correlation b/w weather and health 

## Effect of Bank Access

- **Relaxing financial constraints** increases either or both of $(-b_{i}(r_{t}), b_{i}(r_{t}))$
  - Partial credit modeled as increase in $b_{i}(r_{t})$ only (i.e. can save more)
<!-- loans typically for production only -->
  - Allows agents to smooth consumption; draw down savings to mitigate weather shock
  - Reduces the correlation between weather shocks \( z_{it} \) and health outcomes \( h_{it} \).
\vfill

- **Prediction**: $\text{Prediction: } H_t \downarrow \text{ as } b_i(r_t) \uparrow$
  - Areas with better financial access have smaller effect of $z_{it}$ on health

## Aside: Climate change versus weather

- Climate change is long term, weather is short term

- Studying impacts of weather is easier:
  - Usually use high dimensional fixed effects
  - E.g. with district and state-year FEs, exploit monthly deviations from average
  
- Studying impacts of climate is harder:
  - There is no good counterfactual for climate change

## Deschenes et al. (2013): Weather and Death in India

- **Research question:** how do temperature shocks affect mortality?
  - How does that different between urban/rural areas?
  - What is the role of income as an intermediary?

- **Setting:** Indian districts from 1957-2000

- **Data:** daily gridded weather + annual infant mortality

- **Design:** panel fixed effects
  - Difference-in-differences with rural bank expansion

- **Result:** 1) heat $\uparrow$ mortality, 2) driven by rural populations, 3) credit access mitigates effect by facilitating consumption smoothing

## Data

Rural 

- Yields: annual, district level output/cultivated area (27 crops) (World Bank)

- Ag price index: mean crop price weighted by value in district (World Bank)

- Ag wages: daily wage rate (World Bank)

Urban 

- Manufacturing productivity: output per worker (ASI)

- Urban CPI: NSS survey

- Manufacturing wages: ASI (nominal wage / CPI)

## Empirical Strategy
Flexible specification to model effect of daily temperature:
$$ Y_{dt} = \sum_{j=1}^{11} \theta_{j}TMEAN_{dtj} + \sum_{k} \delta_{k}1\{RAIN_{dt} \text{ in tercile k\}} + \alpha_{d} + \gamma_{t} + \lambda_{r}^{1}t + \lambda_{r}^{2}t^{2} + \epsilon_{dt} $$

- $r$ is a climatic region
- $TMEAN$ is days in year $t$ where: 
  - daily mean temp $<70\degree$, $\geq 97\degree$, or in the nine 3$\degree$ bins in between 
- $\theta_{j}$'s capture non-linearities, despite having annual unit of analysis
- $RAIN$ accounts for correlation b/w temperature and rainfall

## Identifying Assumptions?

<!-- - Impact governed by daily mean alone -->

<!-- - Impact of days' mean temperature on annual mortality is constant within 3$\degree$ intervals -->
<!--   - This is reasonable -->

<!-- - By using number of days in each bin, assume that the sequence of hot and cold days is irrelevant.  -->

<!-- - Annual deviation from district average is random -->

<!-- - Anything else? Reverse causality? -->


## Secondary Approach

- Consolidates same information
- Workhorse model nowadays because more parsimonious
$$Y_{dt} = \beta CDD 80_{dt} + \sum_{k=1}^{3} \delta_{k}1\{RAIN_{dt} \text{ in tercile k\}} + \alpha_{d} + \gamma_{t} + \lambda_{r}^{1}t + \lambda_{r}^{2}t^{2} + \epsilon_{dt}$$

- where CDD80 is cumulative degree days in district that exceeded 80$\degree$.
  e.g. if two days over 80, one at 82 and other at 84, then CDD=2+4=6
- All bins $j < 80\degree$ are zero 
  - The three above $80 \degree$ restricted to be linearly increasing in average temp  
- Assumption: days when mean daily temp below 80 irrelevant for determining Y

## Results: Yields
Test of indirect channel through labor productivity, $a_{i}(z_{it})$
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "55%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/des2013_fig1.png",
                        auto_pdf = TRUE
                        )

```
A single day with temperature over 85$\degree$ reduces yields by 0.5\%

## Results: Wages
Test of indirect channel through $a_{i}(z_{it})$
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "55%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/des2013_fig2.png",
                        auto_pdf = TRUE
                        )

```
A single day with temperature over 85$\degree$ reduces ag wages by 0.25-0.5\%

## Results: Urban vs. Rural
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "85%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/des2013_fig5.png",
                        auto_pdf = TRUE
                        )

```
- This corroborates Prediction 1 from model!

## Results: by Growing Season
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "76%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/des2013_fig4.png",
                        auto_pdf = TRUE
                        )

```
- This is also consistent with Prediction 1 from model!

## Results: Mortality
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "70%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/des2013_fig3.png",
                        auto_pdf = TRUE
                        )

```
Rural: years with more days with exceeding 87$\degree$ have 0.5\% higher mortality 

## Bank Access

- Prediction 2: bank access enables consumption smoothing
  - Relaxing borrowing/saving constraints should mute relationship
\vfill
- Test: Between 1977-1990, central bank extends banking to the poor
  - Existing banked (urban) areas must open 4 branches in unbanked (rural) areas
\vfill
- Method: difference in differences
  - Compare mortality in hot and cold places before and after bank access

## Results: Credit Access
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "85%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/des2013_fig6.png",
                        auto_pdf = TRUE
                        )

```
- Effect of moderate and hot days on mortality is moderated by bank access
- Bank branches more than offset weather-death relationship

**Takeaway: credit enables poor to withstand environmental shocks**

## Discussion

- Aggregate data entangles ambient means and experienced exposure
- What exactly makes rural areas more vulnerable?
- How much of the weather-death effect is direct vs. indirect?
  - Empirically, we did not identify model parameters separately
- Will weather shock be enough to affect remaining lifetime income?

**Way Forward**

- Bank access represents anything that relaxes budget constraint 
- Policy could have been subsidies on $q_{it}$, encouraging seasonal migration, etc.
  - Opean areas for research

##
\centering
\huge Back to the main question

## Back to the question
- How does environmental quality affect economic development?
\vfill
- Most reduced form studies identify $\frac{\partial h}{\partial e}$
\vfill
- Recall, this is an input into estimating benefits (MTWP) of $\uparrow e$
  - which we need to know in order to set optimal environmental policy
  - optimal in the sense that agents decisions are first-best
\vfill
- Is physical health the only thing that matters for estimating $\frac{\partial h}{\partial e}$?

## Chen et al. (2024): Air Pollution and Mental Health

- **Research question:** What is the impact of air pollution on mental health?
  - Direct mechanism: oxidative stress
  - Indirect mechanism: labor productivity, income, physical activity

- **Setting:** China in 2014-2015

- **Data:** Kessler Physchological Distress Scale
  - PM2.5 from China National Environmental Monitoring Center

- **Design:** instrument pollution with thermal inversions

- **Results:** Pollution $\uparrow$ depression
  - Mechanisms: lower exercise, worse physical health
  - Total cost = \$22.88 billion, similar to mortality costs!
  
## Background

- Why do we care about mental health?
  - Poverty associated with life satisfaction and **stress**
  - These mechanisms affect productivity, which perpetuates poverty
  - See Haushofer and Salicath (2023); Haushofer and Fehr (2014)
\vfill
- Frontier in development economics: RCTs on income and stress
  - Cash transfers $\uparrow$ subjective well-being (McGuire et al., 2022)
  - Cash transfers $\downarrow$ suicide in Indonesia (Christian et al., 2019)
  - Cash transfers $\downarrow$ depression (Haushofer and Shapiro, 2016)
\vfill
- Toxicology literature: pollution $\uparrow$ oxidative stress (Calderon-Garciduenas et al., 2003; Sørensen et al., 2003; MohanKumar et al., 2008, Power et al., 2015)

## Background
- Theory of change: **pollution $\rightarrow$ stress** $\rightarrow$ productivity
  - Poverty correlated with pollution and stress
  - Direct effect of pollution on productivity (Graff Zivin \& Niedell, 2012; California)
  - Many parts of the chain are open areas for research!

## Air Pollution and Mental Health

What are the empirical challenges?

1. OVB: pollution correlated with economic activity
    - Richer counties more polluted
    - Induces **negative** bias
    - Solution: ?

2. ?

3. ?

4. ?

## Instrumental Variable: Thermal Inversion

```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "90%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/inversion2.png",
                        auto_pdf = TRUE
                        )

```

## Instrumental Variable: Thermal Inversion
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "70%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/inversion.png",
                        auto_pdf = TRUE
                        )

```

## Instrumental Variable: Thermal Inversion

- Air is warmer near surface, becomes cooler at higher altitudes

- Thermal inversion: temperature-altitude gradient is reversed
  - Rare phenomenon
  - Depends on **random** meteorological factors

- Cool air is denser and does not rise

- Pollution becomes trapped near surface

## What do you think?

<!-- - Pros and cons -->

<!-- - Inversions are short lived (last for a few hours) -->
<!--   - cannot estimate impact of sustained exposure -->
<!--   - Is mental health affected by rare short term events? -->

<!-- - correlated with weather, which can have direct effect on mental health -->
<!--   - control for ground level temperature/rain -->

<!-- - If mechanism is through chronic condition, then TI may isolate wrong type of variation -->

## Data

- Mental health: 2014 China Family Panel Study Survey
  - 16,000 rural and 13,000 urban residents (162 counties out of 1,355)
  - Six questions about mental health on 0-24 Scale
  - Depression, nervousness, restlessness, hopelessness, effort, worthlessness
\vfill
- Pollution: PM2.5
  - Webscraped
  - Validate with satellite data?
\vfill
- Thermal inversion (MERRA-2 Air Temperature Layers)
  - Inversion = 1 if temperature of first layer (110m) $<$ second layer (320m)
  - IV: number of inversions in month prior to each interview

## Inversions and Pollution
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "65%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/oliva2024_fig1.png",
                        auto_pdf = TRUE
                        )

```
Do thermal inversions seem rare and random?

## Empirical Strategy

First Stage:

$$ P_{ict} = \alpha_{0} + \alpha_{1}I_{ict} + f(W_{ict}) + \gamma_{c} + \sigma_{t} + \mu_{ict} $$

Second Stage:
$$ H_{ict} = \beta_{0} + \beta_{1}P_{ict} + f(W_{ict}) + \gamma_{c} + \sigma_{t} + \epsilon_{ict} $$

- $P_{ict}$ is PM2.5 in month prior to interview
- $I_{ict}$ is instrument, number of thermal inversions in past month
- $H_{ict}$ is total mental illness score (0-24)
- $f(W_{ict})$ are flexible weather controls
- $\sigma_{t}$ are year-month FEs (**important**!)

## Main Results: Mental Health
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "90%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/oliva2024_fig2.png",
                        auto_pdf = TRUE
                        )

```
- 1 $\mu/m^{3}$ PM2.5 $\uparrow$ mental illness by 0.1164 units (4\% of mean; 0.4$\sigma$)
- 1 $\mu/m^{3}$ PM2.5 $\uparrow$ prob. of several mental illness by 0.42pp (9.5\% of mean)

## Main Results: Symptoms
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "90%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/oliva2024_fig3.png",
                        auto_pdf = TRUE
                        )

```
- Not driven by any one particular symptom

## Main Results: Persistence
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "90%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/oliva2024_fig4.png",
                        auto_pdf = TRUE
                        )

```
- Effect of pollution exposure on mental health dissipates over time

## Mechanisms
- Focus on indirect effects (physical health, productivity, behavior)

```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "90%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/oliva2024_fig5.png",
                        auto_pdf = TRUE
                        )

```
- Driven by reduction in physical exercise, not other illness
- Very small impact compared to pollution effect (0.002$\sigma$ vs. 0.4$\sigma$)

## Heterogeneity
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "60%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/oliva2024_fig6.png",
                        auto_pdf = TRUE
                        )

```
- Why are effects larger for men?

## Economic Cost? Back of the Envelope

- 1$\sigma$ $\uparrow$ in PM2.5 leads to 9.5\% $\uparrow$ in prob. of severe mental illness
\vfill
  - About 0.095 $\times$ 1.4 billion = 13.3 million additional people 
- Annual cost per patient of mental illness in China is \$USD 3,665 (Xu et. al 2016)
\vfill
- Assume 8\% of patients with mental illness get treated (Phillips et al., 2009)
\vfill
- **Economic cost of mental illness:** $13.3 \times 3665 \times 0.008 \times 12 = 46.7$ billion USD
\vfill
- Mortality cost of $1\sigma$ PM2.5 is \$USD 30 billion in USA (Deryugina et al., 2016)

\vfill
**Issues with the back of the envelope calculation?**
<!-- 1) for severe mental illness; 2) what about non-medial costs? productivity
3) hard to interpret in sd. do a counterfactual with policy-relevant change in pollution -->


## Discussion

## Next week

- productivity and direct income effects

##
\centering
\huge Lecture 5

\large How does the environment affect economic development?

## Housekeeping

- See last week's slides on pollution and mental health (Chen, Oliva \& Zhang, 2024)
\vfill
- Thursday: do your best for the reading (we haven't learned WTP yet)
\vfill
- Thursday: please print the slides and bring to class!
\vfill
- Research Proposal: first draft due two weeks from Thursday
\vfill
- Start thinking about your replication

## Today

- **Guiding question:** how does the environment affect economic development?

\begin{align}
MWTP_{e} = \frac{\lambda_{e}}{\lambda_{y}} 
 = \frac{1}{\lambda_{y}} \left(\frac{\partial u}{\partial e} + \frac{\partial u}{\partial h}\frac{\partial h}{\partial e}\right) + {\color{red}\frac{\partial \Delta y}{\partial e} + \frac{\partial \Delta y}{\partial h}\frac{\partial h}{\partial e}} \notag
\end{align}

\vfill
- Today's focus: productivity and income
  - Indirect: worker productivity (lots of research)
  - Direct: income (very little research)

## Environment, Productivity, and Income

- Temperature is neg. associated w/ aggregate output (Dell, Jones, and Olken 2012)
  - Specifically in poor countries
  - Driven by agriculture and manufacturing sectors
  - Level effects (e.g. output/crop yield) vs. growth rate effects
  
\vfill
- Evidence on direct channel: high temperatures associated w/ lower crop yields
  - Mendelsohn and Dinar (1999), Auffhammer et al. (2006), Schlenker and Roberts (2009),
  Lobell et al. (2011), and Gupta et al. (2017)
  - Very little evidence on other direct channels (e.g. ecotourism)
\vfill
- Evidence on indirect channel via productivity: see reading list

\vfill
- Mechanisms: crop stress, human stress/productivity, conflict, institutions?
  - Open and policy-relevant research area

## Somanathan et al. (2021): Temperature, Productivity, and Labor

- **Research question:** What is the effect of temperature on output?
  - How does heat stress mediate the temperature-output relationship
\vfill
- **Setting:** India's manufacturing sector (cloth weaving, sewing, steel)
\vfill
- **Data:** High-frequency (15 year panel, daily)
  - Worker-level, sewing line level, team level, plant-year, district-year
\vfill
- **Design:** panel fixed effects
\vfill
- **Result:** 1) heat $\downarrow$ plant-level annual output, 2) driven by output elasticity of labor

## Background

- Workers should respond quickly when working in uncomfy temperatures
  - Heat impacts on labor can be identified from daily or aggregated data
  - Distinguishes heat stress from alternative mechanisms which materialized over long time scales (i.e. conflict)
\vfill
- Two channels through which temperatures affect workers
  - Produce less (productivity)
  - Absenteeism
\vfill
- Can we expect lumpy investments in climate control in developing countries?
  - Depends on value-added per worker
  - If too low, then electricity cost not worth it

## Data: Worker + Cloth Weaving

- Worker data
  - Surveyed **output and attendance** from cloth-weaving, garment, and steel firms
  - Three cloth-weaving factories in Surat,
  - Eight garment sewing plants (Delhi, Hyderabad, Chhindwara)
  - Steel plant (one large plant in Central India)
\vfill
- Cloth data (2012): 147 workers
  - Daily meters woven (for each worker)
  - Worker pay slips each day (output + attendance)

## Data: Garment Sewing
- Garment sewing (2012-2013): Line level
  - Lines of 10-20 workers
  - Productivity: hourly output relative to target
  - Daily production from 103 sewing lines
\vfill
- Climate control
  - Phased rollout of cooling equipment
  - 84 of 103 sewing lines in Delhi, of which 74 had cooling

## Data: Rail and Steel Plant
- Supplies rectangular blocks of steel to Indian Railways
- Three shifts per day
- Workers assigned to one of three teams, which rotate across shifts
- Data:
  - Team and number of blocks for each shift during 1999-2008
  - 9,172 shifts over 3,337 working days
  - Team-level absences over 857 days for 2000-2003

## Garment factory and cloth weaving
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "70%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/anant2021_fig7.png",
                        auto_pdf = TRUE
                        )

```

## Rail/Steel Mill
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "70%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/anant2021_fig8.png",
                        auto_pdf = TRUE
                        )

```

## Data: Panel
- Annual Survey of Industries (1999-2013)
  - Census of large manufacturing plants ($>$ 100 workers)
  - Random sample of small plants
  - Output, fixed assets, input expenses, employment
  - Unbalanced panel of 58,377 plants
\vfill
- District Manufacturing GDP (1998-2009)
  - Planning Commission of India
  - ASI plants + unregistered plants
\vfill
- Weather: public weather stations + gridded data (IMD)
  - Daily maximum at closest station to factory
  - Gridded data for panel data

<!-- ## Data: Diamond Firms -->
<!-- - Survey 150 diamond cutting plants in Surat -->
<!--   - Random sample from 500 registered units -->
<!-- - Five operations: 1) grading 2) marking, 3) rounding, 4) cutting, 5) polish -->
<!-- - Ask about number of workers and use of AC in each operation -->
<!-- - Ask to rate on 1-5 scale importance of each operation to quality of final output -->
<!--   - Why? -->

## Empirical Strategy
Temperature-output estimated with weekly data as follows:

$$y_{iw} = \alpha_{i} + \gamma_{M} + \gamma_{t} + \sum_{j=2}^{J} \beta_{j}T_{iw}^{j} + \theta R_{iw} + \lambda X_{iw} + \epsilon_{iw}$$

- $y_{iw}$ is physical output by worker $i$ in week $w$
- If absent, output=0 for that day before aggregating to week
- $\alpha_{i}$ is individual FE (what does it account for?)
- $\gamma_{M/t}$ are month/year FEs (what does it account for?)
- $T_{j}$ counts days falling into temperature bin $j$ (captures nonlinearities)

## Identification Assumptions

## Results: Garment Sewing
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "70%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/anant2021_fig1.png",
                        auto_pdf = TRUE
                        )

```
- output is lower in hotter weeks **without climate control**
- Moving from lowest to highest temperature bin $\downarrow$ productivity by 8\%
- Note: these are 90\% CI's!

## Results: Cloth and Steel
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "80%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/anant2021_fig2.png",
                        auto_pdf = TRUE
                        )

```

## Lagged Effects
- Previously, $\beta_{j}$ combined productivity and absenteeism channel
- Now we use daily data
$$y_{id} = \alpha_{i} + \gamma_{M} + \gamma_{t} + \sum_{j} \beta_{j}T_{id}^{j} + \sum_{j} \omega_{j}L_{id}^{j} + \theta R_{id} + \lambda X_{id} + \epsilon_{id} $$
- where $T^{j}$ is now an indicator for day $d$ falling into temperature bin $j$
- $X_{id}$ includes day-of-week FE
- $L_{id}^{j}$ counts number of days in bin $j$ during six preceding days

## Discussion

- Is lag specification plausible?
  - Is six days a plausible window to detect cumulative exposure?

- What are the identification assumptions?

- What are the threats?

## Lagged Effects: Productivity
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "70%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/anant2021_fig3.png",
                        auto_pdf = TRUE
                        )

```
- Lower output on hot days in sites w/o climate control
- Weavers: additional day $>35\degree$ in past week $\downarrow$ daily output by 3\%

## Lagged Effects: Absenteeism
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "70%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/anant2021_fig4.png",
                        auto_pdf = TRUE
                        )

```
- How do you interpret these results?
- new research idea: effect of paid leave on temperature-productivity relation

## Annual Survey of Industries: 15-year panel results
$$y_{it} = \alpha_{i} + \gamma_{t} + \sum_{j=2}^{5}\beta_{j}T_{it}^{j} + \theta R_{it} + \epsilon_{it}$$

outcome is log value of plant $i$ output in year $t$

```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "95%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/anant2021_fig5.png",
                        auto_pdf = TRUE
                        )

```

## Recap of Results

- Daily worker data: workers are less productive on hot days
\vfill
- Daily worker data: (kind of) more absenteeism on hot days
\vfill
- Plant panel: **aggregate** output is lower on hot days
  - This is the object we are interested in
  - Remember $\frac{\partial \Delta y}{\partial e}$ and $\frac{\partial \Delta y}{\partial h}\frac{\partial h}{\partial e}$
\vfill

- **How much of the temperature-induced $\Delta Y$ is due to $\Delta MP_{L}$?**
  - As opposed to changes in capital intensity
  - No data on climate control in ASI data

## Lets make a Cobb-Douglas model

Consider a production function:
$$ Y = A \cdot K^{\omega} \cdot L^{\beta}$$

Now modify by letting elasticities be a function of temperature, $T$
$$Y = A(T) \cdot K^{\omega(T)} \cdot L^{\beta(T)}$$

where linearity is assumed:

\begin{align*}
A(T) &= A_{0} + \sum a_{j}T_{j} \\
\omega(T) &= \omega_{0} + \sum \omega_{j}T_{j} \\
\beta(T) &= \beta_{0} + \sum \beta_{j}T_{j}
\end{align*}

## Estimating Equation

- Log-linearizing, we get Equation (7) from the paper:

$$ y = \alpha_{0} + \sum_{j=2}^{5} \alpha_{j}T^{j} + \omega_{0} \cdot k + \sum_{j=2}^{5} \omega_{j}T^{j}k + \beta_{0} \cdot l + \sum_{j=2}^{5} \beta_{j}T^{j}l $$

- Data:
  - k: opening value of equipment and machinery at start of financial year
  - l: number of full-time workers

## Results
- Temperature effects on output elasticity of L is negative and significant
  - i.e. temperature makes labor less productive at higher temperatures

- Temperature effects on output elasticity of K are (weakly) positive

- Controlling for interactions, residual temperature-output effect is zero

- Since overall temperature-output relationship is negative, temperature-induced declines in labor productivity drive the effect

## Results
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "55%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/anant2021_fig6.png",
                        auto_pdf = TRUE
                        )

```

## Discussion

- Identification
- Did they do everything they could with the data?
- Alternative explanations

## Back to the model
\begin{align}
MWTP_{e} = \frac{\lambda_{e}}{\lambda_{y}}
 = \frac{1}{\lambda_{y}} \left(\frac{\partial u}{\partial e} + \frac{\partial u}{\partial h}\frac{\partial h}{\partial e}\right) + {\color{red}\frac{\partial \Delta y}{\partial e}} + \frac{\partial \Delta y}{\partial h}\frac{\partial h}{\partial e} \notag
\end{align}

- Most research in environment/development economics seeks to identify $\frac{\partial h}{\partial e}$
  - At this point, $\frac{\partial h}{\partial e}>0$ is unambiguous
  - Focus is on **magnitudes**

\vfill
- What about indirect effects via productivity ($\frac{\partial \Delta y}{\partial h}\frac{\partial h}{\partial e}$)?
  - This is the research frontier
\vfill
- Next: what about $\frac{\partial \Delta y}{\partial e}$, the direct effect of environment on incomes
  - Almost no research on this

## How might the environment **directly** affect incomes?

**Conceptual Framework**

- Assume landholders vary in distance to a central market

- Each chooses land and labor to maximize farm profits

- New PA imposed that overlaps existing productive land and restricts land use

- **Assuming initially efficient allocation, what happens to net welfare?**

- **What if land rents increase outside the PA?**

## Background: Conservation and Development

- On other hand, PAs may generate tourism income and other positive spillovers

- Surprisingly little evidence on income channel

- Main challenge: non-random placement of PAs


## Sims (2010): Conservation and Development: Evidence from Thailand

- **Research Question:** How does PA policy affect local poverty outcomes?
  - What are the mechanisms?

- **Setting:** Thailand in 2000

- **Data:** Household consumption surveys + PA locations

- **Empirical Strategy:** Instrument PA location with priority watershed status

- **Results:** PAs increase consumption and lower poverty
  - Higher tourism is the main mechanism

## Data

- Wildlife sanctuaries (N=31) and national parks (N=57)
  - Source: World Database on Protected Areas
  - Includes year of establishment
\vfill
- Poverty Mapping
  - Model hh income in 2000 Thai Economic Survey as a function of hh features + assets
  - Use coefficient to predict income of census households
\vfill
- Poverty Measure
$$ FGT(\alpha) = \frac{1}{n}\sum_{y_{i}<p}\left( \frac{p-y_{i}}{p} \right)^{\alpha} $$
where $p$ is poverty line and $y_{i}$ is income

## Empirical Strategy: Ordinary Least Squares

$$y_{ij} = \beta_{1} PA_{ij} + \beta_{2}G_{ij} + \beta_{3}X_{ij} + \alpha_{j} + \epsilon_{j}$$

- $y_{ij}$ is socioeconomic outcome for locality $i$ in district $j$
- $PA_{ij}$ is share of locality land designated as PA prior to 2000
- $G_{ij}$ are fixed geographic controls (slope, elevation, water, etc.)
- $X_{ij}$ is a measure of pre-treatment (historic forest cover, distance to major roads)
- $\alpha_{i}$ are district fixed effects
\vfill

**What variation is exploited?**

**What are the identification assumptions?**

**What are potential omitted characteristics?**

<!-- talk about spatial spillovers -->

## OLS Results
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/sims2010_fig1.png",
                        auto_pdf = TRUE
                        )

```
- Median land share protect is 1/3
- Expected $\uparrow$ in consumption is about 4.5\% (divide by coeff. by 3)

## Instrumental Variables Solution

- Instrument PA placement with **priority watershed status**
\vfill
- Upper watersheds protected to improve **downstream** water quality
  - Watershed protection (PA) not driven by local (headwater) economic needs
\vfill
- First Stage:
 $$ PA_{ij} = \pi_{1}Z_{ij} + \pi_{2}G_{ij} + \pi_{3}X_{ij} + \pi_{4}W_{ij} + \delta_{j} + u_{ij} $$

- $Z_{ij}$ includes a) $i$ overlaps major tributary, b) distance to nearest major tributary .
\vfill
- $\pi_{1}$ picks up variation in PA driven by differences in downstream characteristics

## Watershed Instrument
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "60%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/sims2010_fig2.png",
                        auto_pdf = TRUE
                        )

```

## Discussion
\centering
Is the IV plausible? Why or why not?

## IV Results
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "90%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/sims2010_fig3.png",
                        auto_pdf = TRUE
                        )

```
- Signs agree with OLS, but magnitudes are more than 2x as large
- Consistent with idea that OLS negatively biased

## Mechanisms: Is tourism responsible?

- PAs restricts land use, so why does income increase?
\vfill
- Test 1: Test whether impacts in national parks $>$ in wildlife sanctuaries
  - Idea: NP's encourage tourism whereas sanctuaries only permit ecotourism/research
\vfill
- Test 2: Back-of envelope calculation
  - if we know costs/benefits of tourism, is it enough to explain net $\uparrow$ in consumption?
\vfill
- Test 3: Are benefits higher where tourism is higher
  - Park statistics show that visitation is highest near major cities

## Test 1: National parks vs. Wildlife Sanctuaries
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "90%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/sims2010_fig4.png",
                        auto_pdf = TRUE
                        )

```
- Effects larger where tourism is allowed (NPs)
- Difference in coefficients not statistically significant

## Test 2: Back of the envelope calculation

- Assume income is agricultural and reduced proportionally to land lost
- Assume credit constraints (no savings) $\rightarrow$ income losses = consumption losses
- Consumption cost of PAs is 11\% of income (forest cover coeff. - see paper)
- To achieve 5\% *net* $\uparrow$ in consumption, tourism must generate 16\% consumption
  - Equivalent to 233 Baht per month (0.16 $\times$ mean consumption)
- 10.81 mil. tourists in 2020 and 172,978 hh in PA localities = 61 tourists/hh
- Each tourist would need to contribute: (233*12)/61 = 46 Baht (\$USD 1.20) of net income to generate observed increases in consumption

**Is this reasonable?**

## Test 3: Heterogeneity by Distance to City
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/sims2010_fig5a.png",
                        auto_pdf = TRUE
                        )

```
- Most tourists near city, but effects concentrated at intermediate distance
- Why?

## Test 3: Heterogeneity by Distance to City
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/sims2010_fig5b.png",
                        auto_pdf = TRUE
                        )

```

## Discussion
- What do you think of the paper?
- Do you believe the instrument?
- What are some alternative explanations?